Backward Population Synthesis: Mapping the Evolutionary History of Gravitational-Wave Progenitors

### 2023-02-14 src/tex/ms.tex
adding in discussion about redshift, van Son ref, and remove some instances of "we find"

We find that the progenitor properties and even the ability to produce a
GW150914-like merger event is strongly correlated with hyperparameter settings,
consistent with results from prior forward modeling studies cited above;
different assumptions about the black hole masses at merger in GW150914 imply
wildly different formation channels and ZAMS progenitor masses for this event
and some nearby combinations of ZAMS masses are unable, with any reasonable
hyperparameter settings, to produce a DCO merger like GW150914.  We further
exhibit preliminary results of an ongoing analysis over the entire GWTC-3
catalog \citep{GWTC-3} that suggest that the accretion efficiency in stable mass
transfer may depend non-trivially on the primary black hole mass in merging
systems; our methodology allows for a systematic study of such dependencies,
which we will explore more fully in future work.

@@

We show a portion of the joint posterior of the progenitor parameters and
hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. Each panel
is colored based on the comparison between progenitor parameters (blue),
hyperparameters (orange), or a mix of the two (green). We find that the choice
for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not
unexpected since \textbf{the \citet{Fryer2012} delayed model can significantly reduce} the
natal kick strength at BH formation based on the amount of fallback onto the
proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive
enough that the fallback reduces the natal kick to zero. Because of this we do
not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}. Similarly, we do
not include the ZAMS orbital period and eccentricity, which are correlated with
one another, but not strongly with the other progenitor parameters or
hyperparameters. We find strong correlations between the ZAMS masses, as well as
strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
$f_{\rm{acc}}$ and $\alpha$. We show scatter plots of each of these combinations
in the upper-right inset of the figure.

@@

Finally, we find that $\alpha$ and $f_{\rm acc}$ are largely uncorrelated,
though there exist independent trends in each hyper parameter. This is not
totally unexpected since the physical processes which are described by each
hyperparameter, i.e. stable Roche-lobe overflow and common envelope evolution
are two independent channels. Generally, we find that GW150914-like BBHs tend to
prefer larger accretion efficiencies and have common envelope ejection
efficiencies peaking near $\alpha=1$. One shortcoming of our method
implementation is the application of a single hyper parameter for
$f_{\mathrm{acc}}$ and $\alpha$ for each binary, while both the primary and the
secondary star could in principle be defined by their own hyperparameters that
prescribe the outcomes for when each star fills it's Roche lobe. We reserve full
treatment of this for future work but note that this improvement could reveal
correlations between hyperparameters that are not present in our current
analysis.

>>>>

We find that the progenitor properties and even the ability to produce a
GW150914-like merger event is strongly correlated with hyperparameter settings,
\textbf{consistent with results from prior forward modeling studies cited above};
different assumptions about the black hole masses at merger in GW150914 imply
wildly different formation channels and ZAMS progenitor masses for this event
and some nearby combinations of ZAMS masses are unable, with any reasonable
hyperparameter settings, to produce a DCO merger like GW150914.  We further
exhibit preliminary results of an ongoing analysis over the entire GWTC-3
catalog \citep{GWTC-3} that suggest that the accretion efficiency in stable mass
transfer may depend non-trivially on the primary black hole mass in merging
systems; our methodology allows for a systematic study of such dependencies,
which we will explore more fully in future work.

@@

We show a portion of the joint posterior of the progenitor parameters and
hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. Each panel
is colored based on the comparison between progenitor parameters (blue),
hyperparameters (orange), or a mix of the two (green). \textbf{The choice
for $\sigma$ does not affect the formation of GW150914-like BBHs.} This is not
unexpected since \textbf{the \citet{Fryer2012} delayed model can significantly reduce} the
natal kick strength at BH formation based on the amount of fallback onto the
proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive
enough that the fallback reduces the natal kick to zero. Because of this we do
not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}. Similarly, we do
not include the ZAMS orbital period and eccentricity, which are correlated with
one another, but not strongly with the other progenitor parameters or
hyperparameters.

\textbf{The formation redshift distribution is slightly correlated with the $q_{\rm crit, 3}$ and $f_{acc}$ hyperparameters, though the GW150914-like progenitors generally prefer formation redshifts of $z\sim0.1$ (delay times of $\lesssim0.1\,\rm{Gyr}$). This correlation is due to the preference for longer delay times between ZAMS formation and merger for progenitors which undergo stable mass transfer at Roche overflow instead of a common envelope. This result is in direct agreement with \citet{vanSon2022a} who also find that the common envelope evolution channel produces BBH mergers with shorter delay times than the stable Roche-overflow channel.}

\textbf{The joint posterior exhibits} strong correlations between the ZAMS masses, as well as
strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
$f_{\rm{acc}}$ and $\alpha$. We show scatter plots of each of these combinations
in the upper-right inset of the figure.

@@

Finally, $\alpha$ and $f_{\rm acc}$ are largely uncorrelated,
though there exist independent trends in each hyper parameter. This is not
totally unexpected since the physical processes which are described by each
hyperparameter, i.e. stable Roche-lobe overflow and common envelope evolution
are two independent channels. Generally, we find that GW150914-like BBHs tend to
prefer larger accretion efficiencies and have common envelope ejection
efficiencies peaking near $\alpha=1$. One shortcoming of our method
implementation is the application of a single hyper parameter for
$f_{\mathrm{acc}}$ and $\alpha$ for each binary, while both the primary and the
secondary star could in principle be defined by their own hyperparameters that
prescribe the outcomes for when each star fills it's Roche lobe. We reserve full
treatment of this for future work but note that this improvement could reveal
correlations between hyperparameters that are not present in our current
analysis.
